---
title: " Practical Machine Learning Project - Coursera"
author: "rChul"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Packages and set.seed
We're doing 3 steps here

* caret package to do the training phase
* parallel and doParallel package for the multithread performance in R.
* Also set.seed to 333

```{r packages, warning = FALSE, message = FALSE}
library(caret)
set.seed(333)
library(parallel)
library(doParallel)
```

## Parallel Processing Initiation
The code to operate the Parallel package (set CPU counts)

```{r parallel}
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)
```

## Loading Data sets
Loading data that provided by the Assignment

```{r datag}
data <- read.csv("pml-training.csv")
testdata <- read.csv("pml-testing.csv")
```

## Peeking the Data and Cleaning up
As stated in the assignment, We will have to select the variables that reflect the activities of accelometers strapped
with belt ,arm ,forearm ,dumbbell. Naturally,We have to look at the columns name that reflect fore-mentioned activities and elimiate unimportant variables.

```{r data, eval = FALSE}
head(data)
colnames(data)
```

We know that accel represent the Accelerometer so we filter that colnames.
Further, I want to predict using raw sensors raw data so I eliminate total columns and also the var columns since they contain a lot of NA values

```{r grep}
grep1 <- grep("belt|arm|forearm|dumbbell",names(data),value = TRUE)
grep2 <- grep("accel",grep1,value = TRUE)
grep3 <- grep("^total|^var",grep2,value = TRUE,invert = TRUE)
```

## Slicing Data
Since we already have seperate test data, I decided to slice the train data set to 2 pieces which is training and validation data set.

```{r datapart}
inTrain <- createDataPartition(y = data$classe, p = 0.7, list = FALSE)
train = data[inTrain,c(grep3,"classe")]
validate = data[-inTrain,c(grep3,"classe")]
test = testdata[,c(grep3)]
```

## Fitting Control
We have to setip the fit control to set the K-Fold Value (5) as well as Enable Parallel

```{r control}
fitControl <- trainControl(method = "cv",number = 5,allowParallel = TRUE)
```

## Model Fit
Now I train the data using Random forest method, because in order for the certain class of action to complete 
there is a conditions of variables which are required to be met. Futhermore, The results of the prediction are in 
multi-level factor variable which can suit well with decision tree-like. Finally,In order for an accuracy and 
reducing overfit chance. The Random forest seem to be the sensible model.

```{r fit}
modFit1 <- train(classe ~ ., data = train, method = "rf", trControl = fitControl)
modFit1$finalModel
```

As we can see OOB estimation error rate is 5.55% so we can expect out of sample error of around 5%

## Fit on Validation Set
Just to make sure we fit the model to validation set agian.
```{r vali}
validation <- predict(modFit1,validate)
```

##Validation Set Accuracy
The accuracy came at around 95% again which seem to be a consistant and sensible result.
```{r valich}
confusionMatrix(validate$classe,validation)
```


## Fit to Test Data Set & Predict Result
The moment of truth, I fit the model to the test data set and get the prediction.

(Note : If you put this result in the last test you'll get 19/20 right (95%))
```{r result}
testans <- data.frame(predict(modFit1,test))
testans
```

## Disbale Parallel Processing
Needed to set R back to normal.
```{r end}
stopCluster(cluster)
registerDoSEQ()
```